Validate model

### 2023-12-11 Code/06_CreateTablesandPlots.Rmd
Added Uno's C-index

    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/all.centers.", model, ".", type, ".png"))
    out <- PredictionTools::val.surv.mi(p,S,g=5,main="All centers",
                                        time=NULL,lim=c(0,1),dist=TRUE)
    grDevices::dev.off()

    final.full.table <- eval(parse(text=paste0("final.", type, ".table.", model)))
    final.full.table["C-index", "HR"] <- sprintf("%.2f", out$cindex-optimism)
    final.full.table["C-index", "CI"] <- paste0("[", sprintf("%.2f", out$cindex.lower-optimism),
                                                   "; ", sprintf("%.2f", out$cindex.upper-optimism), "]")

@@

# Add the C-index for calibrated MSM model to Table for recurrence
```{r, eval=TRUE}
for (type in c("full", "BS")){
  f.mi <- eval(parse(text=paste0("f.mi.", type, ".Rec.5")))
  cindex.MSM.cal <- rep(0, m)
  cindex.se.MSM.cal <- rep(0, m)
  for (i in 1:m){
    rc <- Hmisc::rcorr.cens(-predict(f.mi,
                                     newdata=mice::complete(imputed.training.data,i), type="lp"), 
                            S.MSM.5.training)
    cindex.MSM.cal[i] <- rc["C Index"]
    cindex.se.MSM.cal[i] <- rc["S.D."]/2
  }
  cindex.mi.MSM.cal <- Rubin.combine(cindex.MSM.cal, cindex.se.MSM.cal)

  final.full.table <- eval(parse(text=paste0("final.", type, ".table.Rec.5")))
  final.full.table["C-index MSM", "HR"] <- sprintf("%.2f", cindex.mi.MSM.cal$est-optimism)
  final.full.table["C-index MSM", "CI"] <- paste0("[", sprintf("%.2f", cindex.mi.MSM.cal$est+qnorm(.025)*cindex.mi.MSM.cal$se-optimism),
                                                 "; ", sprintf("%.2f", cindex.mi.MSM.cal$est+qnorm(.975)*cindex.mi.MSM.cal$se-optimism), "]")

@@

# set rownames
final.full.table.Rec.5 <- as.data.frame(final.full.table.Rec.5)
row.names(final.full.table.Rec.5) <- c("Positive SN status", "Male",
                                    "Age iqr", "Ulceration", "Location",
                                    "Arm", "Leg", "Trunk", "Head and neck",
                                    "Histology", "SSM", "NM", "ALM", "Other",
                                    "Breslow iqr", "Multiple fields", 
                                    "Total number of negative nodes",
                                    "Total number of positive nodes",
                                    "SN tumour burden iqr", 
                                    "Location metastasis in lymph node",
                                    "Subcap", "Combined", "Parenchymal",
                                    "Multifocal", "Extensive",
                                    rownames(final.full.table.Rec.5)[(nrow(final.full.table.Rec.5)-1):nrow(final.full.table.Rec.5)])
final.full.table.MSM.5.refit <- as.data.frame(final.full.table.MSM.5.refit)
row.names(final.full.table.MSM.5.refit) <- c("Positive SN status", "Male",
                                          "Age", "Ulceration", "Location",
                                          "Arm", "Leg", "Trunk", "Head and neck",
                                          "Histology", "SSM", "NM", "ALM", "Other",
                                          "Breslow", "Multiple fields", 
                                          "Total number of negative nodes",
                                          "Total number of positive nodes",
                                          "SN tumour burden", 
                                          "Location metastasis in lymph node",
                                          "Subcap", "Combined", "Parenchymal",
                                          "Multifocal", "Extensive",
                                          row.names(final.full.table.MSM.5.refit)[nrow(final.full.table.MSM.5.refit)])
final.BS.table.Rec.5 <- as.data.frame(final.BS.table.Rec.5)
row.names(final.BS.table.Rec.5) <- c("SN status",
                                     "Negative SN status",
                                     "Positive SN status", 
                                     "Age", 
                                     "Ulceration",
                                     "Location", 
                                     "Upper limb", 
                                     "Lower limb", 
                                     "Trunk", 
                                     "Head and neck",
                                     "Breslow",
                                     "Breslow (negative)",
                                     "Breslow (positive)",
                                     "SN tumour burden", 
                                   row.names(final.BS.table.Rec.5)[(nrow(final.BS.table.Rec.5)-1):nrow(final.BS.table.Rec.5)])
final.BS.table.MSM.5.refit <- as.data.frame(final.BS.table.MSM.5.refit)
row.names(final.BS.table.MSM.5.refit) <- c("SN status",
                                     "Negative SN status",
                                     "Positive SN status", 
                                     "Age", 
                                     "Ulceration",
                                     "Location", 
                                     "Upper limb", 
                                     "Lower limb", 
                                     "Trunk", 
                                     "Head and neck",
                                     "Breslow",
                                     "Breslow (negative)",
                                     "Breslow (positive)",
                                     "SN tumour burden",  
                                   row.names(final.BS.table.MSM.5.refit)[nrow(final.BS.table.MSM.5.refit)])

>>>>

    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/all.centers.", model, ".", type, ".png"))
    out <- PredictionTools::val.surv.mi(p=p, y=S, g=5, main="All centers", CI.metrics=TRUE,
                                        time=NULL, lim=c(0,1), dist=TRUE)
    grDevices::dev.off()

    final.full.table <- eval(parse(text=paste0("final.", type, ".table.", model)))
    final.full.table["Harrell's C-index", "HR"] <- sprintf("%.2f", out$cindex-optimism)
    final.full.table["Harrell's C-index", "CI"] <- paste0("[", sprintf("%.2f", out$cindex.lower-optimism),
                                                   "; ", sprintf("%.2f", out$cindex.upper-optimism), "]")
    final.full.table["Uno's C-index", "HR"] <- sprintf("%.2f", out$uno.C-optimism)

@@

# Add the C-index for calibrated MSM model to Table for recurrence
```{r, eval=TRUE}
for (type in c("full", "BS")){
  f.mi <- eval(parse(text=paste0("f.mi.", type, ".Rec.5")))
  unoC.MSM <- rep(0, m)
  cindex.MSM.cal <- rep(0, m)
  cindex.se.MSM.cal <- rep(0, m)
  for (i in 1:m){
    lp.i <- predict(f.mi, newdata=mice::complete(imputed.training.data,i), type="lp")
    rc <- Hmisc::rcorr.cens(-lp.i, 
                            S.MSM.5.training)
    cindex.MSM.cal[i] <- rc["C Index"]
    cindex.se.MSM.cal[i] <- rc["S.D."]/2

    unoC.MSM[i] <- survC1::Est.Cval(mydata=cbind(S.MSM.5.training[, 1], 
                                              S.MSM.5.training[, 2], 
                                              lp.i), 
                                 tau=5, nofit=TRUE)$Dhat
  }
  cindex.mi.MSM.cal <- Rubin.combine(cindex.MSM.cal, cindex.se.MSM.cal)

  final.full.table <- eval(parse(text=paste0("final.", type, ".table.Rec.5")))
  final.full.table["Harrell's C-index MSM", "HR"] <- sprintf("%.2f", 
                                                cindex.mi.MSM.cal$est-optimism)
  final.full.table["Harrell's C-index MSM", "CI"] <- paste0("[", 
    sprintf("%.2f", cindex.mi.MSM.cal$est+qnorm(.025)*cindex.mi.MSM.cal$se-optimism),
    "; ", sprintf("%.2f", cindex.mi.MSM.cal$est+qnorm(.975)*cindex.mi.MSM.cal$se-optimism), "]")
  final.full.table["Uno's C-index MSM", "HR"] <- sprintf("%.2f", mean(unoC.MSM)-optimism)

@@

# set rownames
final.full.table.Rec.5 <- as.data.frame(final.full.table.Rec.5)
row.names(final.full.table.Rec.5) <- c("Positive SN status", "Male",
                                    "Age iqr", "Ulceration", "Location",
                                    "Arm", "Leg", "Trunk", "Head and neck",
                                    "Histology", "SSM", "NM", "ALM", "Other",
                                    "Breslow iqr", "Multiple fields", 
                                    "Total number of negative nodes",
                                    "Total number of positive nodes",
                                    "SN tumour burden iqr", 
                                    "Location metastasis in lymph node",
                                    "Subcap", "Combined", "Parenchymal",
                                    "Multifocal", "Extensive",
                                    rownames(final.full.table.Rec.5)[(nrow(final.full.table.Rec.5)-3):nrow(final.full.table.Rec.5)])
final.full.table.MSM.5.refit <- as.data.frame(final.full.table.MSM.5.refit)
row.names(final.full.table.MSM.5.refit) <- c("Positive SN status", "Male",
                                          "Age", "Ulceration", "Location",
                                          "Arm", "Leg", "Trunk", "Head and neck",
                                          "Histology", "SSM", "NM", "ALM", "Other",
                                          "Breslow", "Multiple fields", 
                                          "Total number of negative nodes",
                                          "Total number of positive nodes",
                                          "SN tumour burden", 
                                          "Location metastasis in lymph node",
                                          "Subcap", "Combined", "Parenchymal",
                                          "Multifocal", "Extensive",
                                          row.names(final.full.table.MSM.5.refit)[(nrow(final.full.table.MSM.5.refit)-1):nrow(final.full.table.MSM.5.refit)])
final.BS.table.Rec.5 <- as.data.frame(final.BS.table.Rec.5)
row.names(final.BS.table.Rec.5) <- c("SN status",
                                     "Negative SN status",
                                     "Positive SN status", 
                                     "Age", 
                                     "Ulceration",
                                     "Location", 
                                     "Upper limb", 
                                     "Lower limb", 
                                     "Trunk", 
                                     "Head and neck",
                                     "Breslow",
                                     "Breslow (negative)",
                                     "Breslow (positive)",
                                     "SN tumour burden", 
                                   row.names(final.BS.table.Rec.5)[(nrow(final.BS.table.Rec.5)-3):nrow(final.BS.table.Rec.5)])
final.BS.table.MSM.5.refit <- as.data.frame(final.BS.table.MSM.5.refit)
row.names(final.BS.table.MSM.5.refit) <- c("SN status",
                                     "Negative SN status",
                                     "Positive SN status", 
                                     "Age", 
                                     "Ulceration",
                                     "Location", 
                                     "Upper limb", 
                                     "Lower limb", 
                                     "Trunk", 
                                     "Head and neck",
                                     "Breslow",
                                     "Breslow (negative)",
                                     "Breslow (positive)",
                                     "SN tumour burden",  
                                   row.names(final.BS.table.MSM.5.refit)[(nrow(final.BS.table.MSM.5.refit)-1):nrow(final.BS.table.MSM.5.refit)])

### 2023-12-11 Code/07_CreateCrossValidatedCalibrationPlots.Rmd
Added Uno's C-index

        png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/", model, ".", select.value, ".", type, ".", Centers[j], ".png"))
        out <- PredictionTools::val.surv.mi(p, S, g=5, main=paste("Center", Centers.anonimized[j]),
                                            time=NULL,lim=c(0,1),dist=TRUE)
        grDevices::dev.off()
      }
    }
  }
}

>>>>

        png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/", model, ".", select.value, ".", type, ".", Centers[j], ".png"))
        out <- PredictionTools::val.surv.mi(p, S, g=5, main=paste("Center", Centers.anonimized[j]),
                                            time=NULL,lim=c(0,1),dist=TRUE,
                                            CI.metrics=TRUE)
        grDevices::dev.off()
      }
    }
  }
}

### 2023-12-11 Code/10_AJCC.Rmd
Final commit
Added Uno's C-index

---
title: "Calculate AJCC"
author: "Carolien C.H.M. Maas"
date: "December 2022"
output: html_document
---

@@

# load libraries
library(PredictionTools)  # to plot calibration
library(compareC)         # compare AJCC to model
```
# Obtain C-index of AJCC7 and AJCC8
```{r, eval=TRUE}
calculate.AJCC.7.8 <- function(imputed.data=NULL, type=NULL, S.Rec.5=NULL){
  ####
  #### AJCC 7
  ####
  TNM.7 <- c()
  AJCC.7 <- c()
  cindex.AJCC.7 <- rep(0, m)
  cindex.AJCC.7.se <- rep(0, m)
  cindex.AJCC.7.pos <- rep(0, m)
  cindex.AJCC.7.se.pos <- rep(0, m)
  cindex.AJCC.7.neg <- rep(0, m)
  cindex.AJCC.7.se.neg <- rep(0, m)
  for (i in 1:m){
    imputation.i <- mice::complete(imputed.data,i)
    TNM.7.i <- rep("", nrow(imputation.i))
    AJCC.7.i <- rep(0, nrow(imputation.i))
    for (patient in 1:nrow(imputation.i)){
      breslow.patient <- imputation.i[patient, "Breslow"]
      ulceration.patient <- as.character(imputation.i[patient, "Ulceration"])
      mitosis.patient <- as.character(imputation.i[patient, "Mitosis"])
      nodes.patient <- imputation.i[patient, "Tot_SNs_pos"]

@@

    # save TNM.7 and AJCC.7 for each imputated data set
    TNM.7 <- cbind(TNM.7, TNM.7.i)
    AJCC.7 <- cbind(AJCC.7, AJCC.7.i)

    # compute C-index
    rc <- Hmisc::rcorr.cens(-AJCC.7.i, S.Rec.5) # add a minus to revert class order: from worst to best
    cindex.AJCC.7[i]<-rc["C Index"]
    cindex.AJCC.7.se[i]<-rc["S.D."]/2

    # compute C-index for only positive sentinel node patients
    rc.pos <- Hmisc::rcorr.cens(-AJCC.7.i[imputation.i$SNstatus=="Positive"], S.Rec.5[imputation.i$SNstatus=="Positive",])
    cindex.AJCC.7.pos[i]<-rc.pos["C Index"]
    cindex.AJCC.7.se.pos[i]<-rc.pos["S.D."]/2

    # compute C-index for only negative sentinel node patients
    rc.neg <- Hmisc::rcorr.cens(-AJCC.7.i[imputation.i$SNstatus=="Negative"], S.Rec.5[imputation.i$SNstatus=="Negative",])
    cindex.AJCC.7.neg[i]<-rc.neg["C Index"]
    cindex.AJCC.7.se.neg[i]<-rc.neg["S.D."]/2
  }

@@

  # save C-index and confidence interval
  AJCC.7.table <- rbind(c("All patients:", sprintf("%.2f", cindex.mi.AJCC.7$est),
                paste0("[", sprintf("%.2f", cindex.mi.AJCC.7$est+qnorm(.025)*cindex.mi.AJCC.7$se),
                        "; ", sprintf("%.2f", cindex.mi.AJCC.7$est+qnorm(.975)*cindex.mi.AJCC.7$se), "]")),
              c("Negative SN patients:", sprintf("%.2f", cindex.mi.AJCC.7.neg$est),
                    paste0("[", sprintf("%.2f", cindex.mi.AJCC.7.neg$est+qnorm(.025)*cindex.mi.AJCC.7.neg$se),
                           "; ", sprintf("%.2f", cindex.mi.AJCC.7.neg$est+qnorm(.975)*cindex.mi.AJCC.7.neg$se), "]")),
              c("Positive SN patients:", sprintf("%.2f", cindex.mi.AJCC.7.pos$est),
                    paste0("[", sprintf("%.2f", cindex.mi.AJCC.7.pos$est+qnorm(.025)*cindex.mi.AJCC.7.pos$se),
                           "; ", sprintf("%.2f", cindex.mi.AJCC.7.pos$est+qnorm(.975)*cindex.mi.AJCC.7.pos$se), "]")))

  ####
  #### AJCC8
  ####
  TNM.8 <- c()
  AJCC.8 <- c()
  cindex.AJCC.8 <- rep(0, m)
  cindex.AJCC.8.se <- rep(0, m)
  cindex.AJCC.8.pos <- rep(0, m)
  cindex.AJCC.8.se.pos <- rep(0, m)
  cindex.AJCC.8.neg <- rep(0, m)
  cindex.AJCC.8.se.neg <- rep(0, m)
  for (i in 1:m){
    imputation.i <- mice::complete(imputed.data,i)
    TNM.8.i <- rep("", nrow(imputation.i))
    AJCC.8.i <- rep(0, nrow(imputation.i))
    for (patient in 1:nrow(imputation.i)){
      breslow.patient <- imputation.i[patient, "Breslow"]
      ulceration.patient <- as.character(imputation.i[patient, "Ulceration"])
      mitosis.patient <- as.character(imputation.i[patient, "Mitosis"])
      nodes.patient <- imputation.i[patient, "Tot_SNs_pos"]

@@

    # save TNM.8 and AJCC.8 for each imputated data set
    TNM.8 <- cbind(TNM.8, TNM.8.i)
    AJCC.8 <- cbind(AJCC.8, AJCC.8.i)

    # compute C-index
    rc <- Hmisc::rcorr.cens(-AJCC.8.i, S.Rec.5) # add a minus to revert class order: from worst to best
    cindex.AJCC.8[i]<-rc["C Index"]
    cindex.AJCC.8.se[i]<-rc["S.D."]/2

    # compute C-index for only positive sentinel node patients
    rc.pos <- Hmisc::rcorr.cens(-AJCC.8.i[imputation.i$SNstatus=="Positive"], S.Rec.5[imputation.i$SNstatus=="Positive",])
    cindex.AJCC.8.pos[i]<-rc.pos["C Index"]
    cindex.AJCC.8.se.pos[i]<-rc.pos["S.D."]/2

    # compute C-index for only negative sentinel node patients
    rc.neg <- Hmisc::rcorr.cens(-AJCC.8.i[imputation.i$SNstatus=="Negative"], S.Rec.5[imputation.i$SNstatus=="Negative",])
    cindex.AJCC.8.neg[i]<-rc.neg["C Index"]
    cindex.AJCC.8.se.neg[i]<-rc.neg["S.D."]/2
  }
  # combine C-index for each imputed data set
  cindex.mi.AJCC.8<-Rubin.combine(cindex.AJCC.8,cindex.AJCC.8.se)
  cindex.mi.AJCC.8.pos<-Rubin.combine(cindex.AJCC.8.pos,cindex.AJCC.8.se.pos)
  cindex.mi.AJCC.8.neg<-Rubin.combine(cindex.AJCC.8.neg,cindex.AJCC.8.se.neg)

  # save C-index and confidence interval
  AJCC.8.table <- rbind(c("All patients:", sprintf("%.2f", cindex.mi.AJCC.8$est),
                          paste0("[", sprintf("%.2f", cindex.mi.AJCC.8$est+qnorm(.025)*cindex.mi.AJCC.8$se),
                                 "; ", sprintf("%.2f", cindex.mi.AJCC.8$est+qnorm(.975)*cindex.mi.AJCC.8$se), "]")),
                    c("Negative SN patients", sprintf("%.2f", cindex.mi.AJCC.8.neg$est),
                          paste0("[", sprintf("%.2f", cindex.mi.AJCC.8.neg$est+qnorm(.025)*cindex.mi.AJCC.8.neg$se),
                                 "; ", sprintf("%.2f", cindex.mi.AJCC.8.neg$est+qnorm(.975)*cindex.mi.AJCC.8.neg$se), "]")),
                    c("Positive SN patients:", sprintf("%.2f", cindex.mi.AJCC.8.pos$est),
                          paste0("[", sprintf("%.2f", cindex.mi.AJCC.8.pos$est+qnorm(.025)*cindex.mi.AJCC.8.pos$se),
                                 "; ", sprintf("%.2f", cindex.mi.AJCC.8.pos$est+qnorm(.975)*cindex.mi.AJCC.8.pos$se), "]")))

@@

# Write to Excel
```{r, eval=TRUE}
utils::write.table(cbind(train.AJCC$AJCC.7.table, rep("", 3), train.AJCC$AJCC.8.table[,-1]),
              file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/AJCC.train.txt"),
              row.names=FALSE, col.names=FALSE, sep=",")
utils::write.table(cbind(CIndex.val, rep("", 3),
                   validation.AJCC$AJCC.7.table[, -1], rep("", 3), 
                   compare.AJCC.model[, 2:3], rep("", 3),
                   validation.AJCC$AJCC.8.table[,-1], rep("", 3),
                   compare.AJCC.model[, 4:5]),
              file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/AJCC.validation.txt"),
              row.names=FALSE, col.names=FALSE, sep=",")
utils::write.table(cbind(train.AJCC$AJCC.7.8.count, rep("", 3), validation.AJCC$AJCC.7.8.count[,-1]),
              file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/AJCC.7.8.count.txt"),
              row.names=FALSE, col.names=FALSE, sep=",")
```

>>>>

---
title: "Calculate AJCC"
author: "Carolien C.H.M. Maas"
date: "November 2022"
output: html_document
---

@@

# load libraries
library(PredictionTools)  # to plot calibration
library(compareC)         # compare AJCC to model
```
# Obtain C-index of AJCC7 and AJCC8
```{r, eval=TRUE}
calculate.AJCC.7.8 <- function(imputed.data=NULL, type=NULL, S.Rec.5=NULL){
  ####
  #### AJCC 7
  ####
  TNM.7 <- c()
  AJCC.7 <- c()
  unoC.AJCC.7 <- rep(0, m)
  cindex.AJCC.7 <- rep(0, m)
  cindex.AJCC.7.se <- rep(0, m)
  unoC.AJCC.7.pos <- rep(0, m)
  cindex.AJCC.7.pos <- rep(0, m)
  cindex.AJCC.7.se.pos <- rep(0, m)
  unoC.AJCC.7.neg <- rep(0, m)
  cindex.AJCC.7.neg <- rep(0, m)
  cindex.AJCC.7.se.neg <- rep(0, m)
  for (i in 1:m){
    imputation.i <- mice::complete(imputed.data,i)
    TNM.7.i <- rep("", nrow(imputation.i))
    AJCC.7.i <- rep(0, nrow(imputation.i))
    for (patient in 1:nrow(imputation.i)){
      breslow.patient <- imputation.i[patient, "Breslow"]
      ulceration.patient <- as.character(imputation.i[patient, "Ulceration"])
      if (type=="training"){
        mitosis.patient <- as.character(imputation.i[patient, "Mitosis"])
      } else{
        mitosis.patient <- ifelse(imputation.i[patient, "Mitosis"]<1, "Yes", "No")
      }
      nodes.patient <- imputation.i[patient, "Tot_SNs_pos"]

@@

    # save TNM.7 and AJCC.7 for each imputed data set
    TNM.7 <- cbind(TNM.7, TNM.7.i)
    AJCC.7 <- cbind(AJCC.7, AJCC.7.i)

    # compute Harrell's C-index
    rc <- Hmisc::rcorr.cens(-AJCC.7.i, S.Rec.5) # add a minus to revert class order: from worst to best
    cindex.AJCC.7[i]<-rc["C Index"]
    cindex.AJCC.7.se[i]<-rc["S.D."]/2

    # Uno's c-index  
    unoC.AJCC.7[i] <- survC1::Est.Cval(mydata=cbind(S.Rec.5[, 1], S.Rec.5[, 2], 
                                              AJCC.7.i), 
                                 tau=5, nofit=TRUE)$Dhat

    # compute C-index for only positive sentinel node patients
    positives <- imputation.i$SNstatus=="Positive"
    rc.pos <- Hmisc::rcorr.cens(-AJCC.7.i[positives], S.Rec.5[positives,])
    cindex.AJCC.7.pos[i]<-rc.pos["C Index"]
    cindex.AJCC.7.se.pos[i]<-rc.pos["S.D."]/2

    # Uno's c-index  
    unoC.AJCC.7.pos[i] <- survC1::Est.Cval(mydata=cbind(S.Rec.5[positives, 1], 
                                              S.Rec.5[positives, 2], 
                                              AJCC.7.i[positives]), 
                                 tau=5, nofit=TRUE)$Dhat

    # compute C-index for only negative sentinel node patients
    negatives <- imputation.i$SNstatus=="Negative"
    rc.neg <- Hmisc::rcorr.cens(-AJCC.7.i[negatives], S.Rec.5[negatives,])
    cindex.AJCC.7.neg[i]<-rc.neg["C Index"]
    cindex.AJCC.7.se.neg[i]<-rc.neg["S.D."]/2

    # Uno's c-index  
    unoC.AJCC.7.neg[i] <- survC1::Est.Cval(mydata=cbind(S.Rec.5[negatives, 1], 
                                              S.Rec.5[negatives, 2], 
                                              AJCC.7.i[negatives]), 
                                 tau=5, nofit=TRUE)$Dhat
  }

@@

  # save C-index and confidence interval
  AJCC.7.table <- rbind(c("All patients:", sprintf("%.2f", cindex.mi.AJCC.7$est),
                paste0("[", sprintf("%.2f", cindex.mi.AJCC.7$est+qnorm(.025)*cindex.mi.AJCC.7$se),
                        "; ", sprintf("%.2f", cindex.mi.AJCC.7$est+qnorm(.975)*cindex.mi.AJCC.7$se), "]"),
                sprintf("%.2f", mean(unoC.AJCC.7))),
              c("Positive SN patients:", sprintf("%.2f", cindex.mi.AJCC.7.pos$est),
                    paste0("[", sprintf("%.2f", cindex.mi.AJCC.7.pos$est+qnorm(.025)*cindex.mi.AJCC.7.pos$se),
                           "; ", sprintf("%.2f", cindex.mi.AJCC.7.pos$est+qnorm(.975)*cindex.mi.AJCC.7.pos$se), "]"),
                sprintf("%.2f", mean(unoC.AJCC.7.pos))),
              c("Negative SN patients:", sprintf("%.2f", cindex.mi.AJCC.7.neg$est),
                    paste0("[", sprintf("%.2f", cindex.mi.AJCC.7.neg$est+qnorm(.025)*cindex.mi.AJCC.7.neg$se),
                           "; ", sprintf("%.2f", cindex.mi.AJCC.7.neg$est+qnorm(.975)*cindex.mi.AJCC.7.neg$se), "]"),
                sprintf("%.2f", mean(unoC.AJCC.7.neg))))

  ####
  #### AJCC8
  ####
  TNM.8 <- c()
  AJCC.8 <- c()
  unoC.AJCC.8 <- rep(0, m)
  cindex.AJCC.8 <- rep(0, m)
  cindex.AJCC.8.se <- rep(0, m)
  unoC.AJCC.8.pos <- rep(0, m)
  cindex.AJCC.8.pos <- rep(0, m)
  cindex.AJCC.8.se.pos <- rep(0, m)
  unoC.AJCC.8.neg <- rep(0, m)
  cindex.AJCC.8.neg <- rep(0, m)
  cindex.AJCC.8.se.neg <- rep(0, m)
  for (i in 1:m){
    imputation.i <- mice::complete(imputed.data,i)
    TNM.8.i <- rep("", nrow(imputation.i))
    AJCC.8.i <- rep(0, nrow(imputation.i))
    for (patient in 1:nrow(imputation.i)){
      breslow.patient <- imputation.i[patient, "Breslow"]
      ulceration.patient <- as.character(imputation.i[patient, "Ulceration"])
      mitosis.patient <- as.character(imputation.i[patient, "Mitosis"])
      nodes.patient <- imputation.i[patient, "Tot_SNs_pos"]

@@

    # save TNM.8 and AJCC.8 for each imputed data set
    TNM.8 <- cbind(TNM.8, TNM.8.i)
    AJCC.8 <- cbind(AJCC.8, AJCC.8.i)

    # compute Harrell's C-index
    rc <- Hmisc::rcorr.cens(-AJCC.8.i, S.Rec.5) # add a minus to revert class order: from worst to best
    cindex.AJCC.8[i]<-rc["C Index"]
    cindex.AJCC.8.se[i]<-rc["S.D."]/2

    # Uno's c-index  
    unoC.AJCC.8[i] <- survC1::Est.Cval(mydata=cbind(S.Rec.5[, 1], S.Rec.5[, 2], 
                                              AJCC.8.i), 
                                 tau=5, nofit=TRUE)$Dhat

    # compute C-index for only positive sentinel node patients
    positives <- imputation.i$SNstatus=="Positive"
    rc.pos <- Hmisc::rcorr.cens(-AJCC.8.i[positives], S.Rec.5[positives,])
    cindex.AJCC.8.pos[i]<-rc.pos["C Index"]
    cindex.AJCC.8.se.pos[i]<-rc.pos["S.D."]/2
    unoC.AJCC.8.pos[i] <- survC1::Est.Cval(mydata=cbind(S.Rec.5[positives, 1], 
                                                    S.Rec.5[positives, 2], 
                                                    AJCC.8.i[positives]), 
                                 tau=5, nofit=TRUE)$Dhat

    # compute C-index for only negative sentinel node patients
    negatives <- imputation.i$SNstatus=="Negative"
    rc.neg <- Hmisc::rcorr.cens(-AJCC.8.i[negatives], S.Rec.5[negatives,])
    cindex.AJCC.8.neg[i]<-rc.neg["C Index"]
    cindex.AJCC.8.se.neg[i]<-rc.neg["S.D."]/2
    unoC.AJCC.8.neg[i] <- survC1::Est.Cval(mydata=cbind(S.Rec.5[negatives, 1], 
                                                    S.Rec.5[negatives, 2], 
                                                    AJCC.8.i[negatives]), 
                                 tau=5, nofit=TRUE)$Dhat

  }
  # combine C-index for each imputed data set
  cindex.mi.AJCC.8<-Rubin.combine(cindex.AJCC.8,cindex.AJCC.8.se)
  cindex.mi.AJCC.8.pos<-Rubin.combine(cindex.AJCC.8.pos,cindex.AJCC.8.se.pos)
  cindex.mi.AJCC.8.neg<-Rubin.combine(cindex.AJCC.8.neg,cindex.AJCC.8.se.neg)

  # save C-index and confidence interval
  AJCC.8.table <- rbind(c("All patients:", sprintf("%.2f", cindex.mi.AJCC.8$est),
                          paste0("[", sprintf("%.2f", cindex.mi.AJCC.8$est+qnorm(.025)*cindex.mi.AJCC.8$se),
                                 "; ", sprintf("%.2f", cindex.mi.AJCC.8$est+qnorm(.975)*cindex.mi.AJCC.8$se), "]"),
                          sprintf("%.2f", mean(unoC.AJCC.8))),
                    c("Positive SN patients:", sprintf("%.2f", cindex.mi.AJCC.8.pos$est),
                          paste0("[", sprintf("%.2f", cindex.mi.AJCC.8.pos$est+qnorm(.025)*cindex.mi.AJCC.8.pos$se),
                                 "; ", sprintf("%.2f", cindex.mi.AJCC.8.pos$est+qnorm(.975)*cindex.mi.AJCC.8.pos$se), "]"),
                      sprintf("%.2f", mean(unoC.AJCC.8.pos))),
                    c("Negative SN patients", sprintf("%.2f", cindex.mi.AJCC.8.neg$est),
                          paste0("[", sprintf("%.2f", cindex.mi.AJCC.8.neg$est+qnorm(.025)*cindex.mi.AJCC.8.neg$se),
                                 "; ", sprintf("%.2f", cindex.mi.AJCC.8.neg$est+qnorm(.975)*cindex.mi.AJCC.8.neg$se), "]"),
                      sprintf("%.2f", mean(unoC.AJCC.8.neg))))

@@

# Write to Excel
```{r, eval=TRUE}
utils::write.table(cbind(train.AJCC$AJCC.7.table, rep("", 3), train.AJCC$AJCC.8.table[,-1]),
              file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/AJCC.train.txt"),
              row.names=FALSE, col.names=FALSE, sep=",")
utils::write.table(cbind(CIndex.val,
                   validation.AJCC$AJCC.7.table[, -1],
                   compare.AJCC.model[, 2:3],
                   validation.AJCC$AJCC.8.table[,-1],
                   compare.AJCC.model[, 4:5]),
              file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/AJCC.validation.txt"),
              row.names=FALSE, col.names=FALSE, sep=",")
utils::write.table(cbind(train.AJCC$AJCC.7.8.count, rep("", 3), validation.AJCC$AJCC.7.8.count[,-1]),
              file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/AJCC.7.8.count.txt"),
              row.names=FALSE, col.names=FALSE, sep=",")
```

### 2023-12-11 Code/12_Validation.Rmd
Added Uno's C-index

    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/", data.type, ".", outcome, ".All.png"))
    par(mar=rep(0, 4))
    PredictionTools::val.surv.mi(p, S,
                                 g=5, main="All patients", time=NULL, 
                                 lim=c(0,1), dist=TRUE,
                                 optimism.C=ifelse(data.type=="training", optimism, 0))
    grDevices::dev.off()

    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/", data.type, ".", outcome, ".Positive.png"))
    par(mar=rep(0, 4))
    PredictionTools::val.surv.mi(p[data$SNstatus=="Positive",],
                                 S[data$SNstatus=="Positive",],
                                 g=5, main="Patients with positive SN", 
                                 time=NULL, lim=c(0,1), dist=TRUE,
                                 optimism.C=ifelse(data.type=="training", optimism, 0))
    grDevices::dev.off()

    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/", data.type, ".", outcome, ".Negative.png"))
    par(mar=rep(0, 4))
    PredictionTools::val.surv.mi(p[data$SNstatus=="Negative",],
                                 S[data$SNstatus=="Negative",],
                                 g=5, main="Patients with negative SN", 
                                 time=NULL, lim=c(0,1), dist=TRUE,
                                 optimism.C=ifelse(data.type=="training", optimism, 0))
    grDevices::dev.off()

>>>>

    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/", data.type, ".", outcome, ".All.png"))
    par(mar=rep(0, 4))
    PredictionTools::val.surv.mi(p, S,
                                 g=5, main="All patients", time=NULL, 
                                 lim=c(0,1), dist=TRUE, CI.metrics=TRUE,
                                 optimism.C=ifelse(data.type=="training", optimism, 0))
    grDevices::dev.off()

    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/", data.type, ".", outcome, ".Positive.png"))
    par(mar=rep(0, 4))
    PredictionTools::val.surv.mi(p[data$SNstatus=="Positive",],
                                 S[data$SNstatus=="Positive",],
                                 g=5, main="Patients with positive SN", 
                                 time=NULL, lim=c(0,1), dist=TRUE, CI.metrics=TRUE,
                                 optimism.C=ifelse(data.type=="training", optimism, 0))
    grDevices::dev.off()

    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/", data.type, ".", outcome, ".Negative.png"))
    par(mar=rep(0, 4))
    PredictionTools::val.surv.mi(p[data$SNstatus=="Negative",],
                                 S[data$SNstatus=="Negative",],
                                 g=5, main="Patients with negative SN", 
                                 time=NULL, lim=c(0,1), dist=TRUE, CI.metrics=TRUE,
                                 optimism.C=ifelse(data.type=="training", optimism, 0))
    grDevices::dev.off()
